From Reversible Programs to \\ Univalent Universes and Back

### 2017-05-25 upi.tex
more on slides
minor tweaks

\newcommand{\amr}[1]{\fbox{\begin{minipage}{0.8\textwidth}\color{red}{Amr says: {#1}}\end{minipage}}}

@@

\begin{abstract}
  We establish a close connection between a reversible programming language
  based on type isomorphisms and a formally presented univalent universe. The
  correspondence relates combinators witnessing type isomorphisms in the
  programming language to paths in the univalent universe, and combinator
  optimizations in the programming language to 2-paths in the univalent
  universe. The result suggests a simple computational interpretation of paths
  and of univalence in terms of familiar programming constructs whenever the
  universe in question is computable.
\end{abstract}

@@

There is a cottage industry of reversible programming languages, reversible
logic, programming applications of type isomorphisms, etc. This work seems it
should be connected to HoTT and univalence but there aren't any precise
connections or theorems.

We propose a featherweight very specialized version of HoTT: a smaller
collection of types and only reversible functions. No dependent types!  More
precisely, our conjecture:

@@

@@

Recall that $\{\bt\} = \Sigma_{(A:\mathcal{U})} \|A \simeq \bt\|$. We begin by
unpacking the objects, 1-paths, and 2-paths of $\{\bt\}$.

>>>>

\newcommand{\amr}[1]{\fbox{\begin{minipage}{0.8\textwidth}\color{red}{Amr says: {#1}}\end{minipage}}}
\newcommand{\jacques}[1]{\fbox{\begin{minipage}{0.8\textwidth}\color{red}{Jacques says: {#1}}\end{minipage}}}

@@

\begin{abstract}
  We establish a close connection between a reversible programming language
  based on type isomorphisms and a formally presented univalent universe. The
  correspondence relates combinators witnessing type isomorphisms in the
  programming language to paths in the univalent universe; combinator
  optimizations in the programming language to 2-paths in the univalent
  universe. The result suggests a simple computational interpretation of paths
  and of univalence in terms of familiar programming constructs whenever the
  universe in question is computable.
\end{abstract}

@@

There is a cottage industry of reversible programming languages, reversible
logic, programming applications of type isomorphisms, etc. It seems that this
work should be connected to HoTT and univalence, but there aren't any precise
connections or theorems.

We propose a very specialized version of HoTT, which we dub
``featherweight HoTT''; it sports a smaller
collection of types and only reversible functions. No dependent types!  More
precisely, our conjecture:

@@

\jacques{Just to make sure: we're aiming to prove a much much smaller version of
this here, right?  Where in fact the only type is Bool?}

@@

\textbf{Theorem.} For any type $F$, the type $\Sigma_{A:\mathcal{U}} \| F \equiv A \|$ is the base
of a univalent fibration.

Recall that
$\{\bt\} = \Sigma_{(A:\mathcal{U})} \|A \simeq \bt\| = \Sigma_{(A:\mathcal{U})}
\| \mathit{Bool} \equiv A\|$. We begin by unpacking the objects, 1-paths, and 2-paths of
$\{\bt\}$.
